\documentclass[12pt]{letter}
\usepackage[utf8]{inputenc}
\usepackage[margin=2.5cm]{geometry}
\usepackage{hyperref}

\signature{Gang Zhang}
\address{}

\begin{document}

\begin{letter}{Editorial Office\\Classical and Quantum Gravity}

\opening{Dear Editor,}

I am submitting the enclosed manuscript, ``Layered Structural Screening of 4D
Lorentzian Causal Sets: Linear Admissibility, Quadratic Identity, and Historical
Sedimentation,'' for consideration as a full paper in Classical and Quantum
Gravity.

\textbf{Summary.}
The causal set path integral faces the Kleitman--Rothschild entropy catastrophe:
generic finite posets are non-geometric, and their combinatorial entropy
overwhelms any known discrete action.
We present numerical evidence, from a 25-family poset library at scales
$N=10$--$1024$, that robust selection of 4D Lorentzian causal sets requires a
\emph{two-layer screening architecture}:
(1)~a linear admissibility gate via the Benincasa--Dowker action, which filters
non-geometric posets but cannot uniquely select 4D (rank 14/17 alone);
(2)~a zero-parameter quadratic identity functional (Mahalanobis distance from the
Lor4D reference manifold), which uniquely identifies 4D Lorentzian causal sets
from $N\geq 10$ onward.
The two layers are statistically independent and their intersection is Lor4D
alone.
The identity basin deepens systematically with $N$, a process we term
\emph{historical sedimentation}.

\textbf{Relation to prior work.}
This manuscript supersedes and substantially extends a previous submission
(CQG-115402), which was desk-rejected due to the use of linear extension counts
(logH) as an entropy proxy.
The present work completely replaces logH with purely causal-geometric
observables (effective dimension, interval ratio, antichain width), expands the
comparison library from 7 to 25 families including 8 adversarial constructions,
and introduces the two-layer screening framework with zero free parameters.
The core claims are immune to the criticisms raised against the earlier
submission.

\textbf{Scope and novelty.}
To our knowledge, this is the first systematic demonstration that dimension
selection in causal sets requires functionally distinct screening layers at
different algebraic orders, and the first identification of a zero-parameter
structural identity functional for 4D Lorentzian causal sets.
The historical sedimentation interpretation---that geometric identity
progressively locks as the causal structure grows---provides a new perspective on
the entropy catastrophe.

The manuscript contains 7 sections, 6 tables, 5 figures, and 10 references.
It discusses the two-layer screening architecture in detail (Prediction~B) and
provides bridging discussion to four related prediction lines (A, C, D, E) that
independently support the framework.
It has not been submitted elsewhere.

\closing{Sincerely,}

\end{letter}
\end{document}
